### 2025-05-09 Chapters/Chapter2.tex
vinna
Updated pages and figures.

\section{Automatic Weather Station Data}

Automatic weather stations (AWS) have been set up all around the world, including Iceland. AWSs might include a variety of sensors, such as barometers, thermometers, and anemometer along with solar panels (or other power source) and some kind of data logger. IMO provided measurements from AWSs across Iceland. One of the stations used in this study can be seen in Figure \ref{fig:storhofdi}. Not every station had consistently available data. Data from \nStationsMin AWSs were included looking at a period from \startDateVedur and ended in 2023. Of these \nStationsMin stations provided by IMO, \nVedurMin were IMO stations, with the anemometer at 10~m above ground, while the remaining \nVGMin stations were from \href{https://www.vegagerdin.is/}{the Icelandic Road and Coastal Administration (IRCA)}, with the anemometer at 6–7~m above ground \cite{vegagerdin_postur}. The locations of these weather stations are shown in Figure \ref{fig:aws_map}.

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{Figures/storhofdi.jpeg}
  \caption[AWS at Stórhöfði in Vestmannaeyjar, south of Iceland.]{AWS at Stórhöfði in Vestmannaeyjar, south of Iceland. The anemometer can be seen as three pins in the upper-central area of the figure just slightly to the left.}
  \label{fig:storhofdi}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/station_elevations.png}
    \caption[Locations of automatic weather stations in Iceland]{Locations of all 412 stations that were looked at in this study. Most of these were from IMO but over a hundred were from IRCA. IMO anemometers are placed at 10 meters above ground, while IRCA ones are placed at 6--7 meters above ground.}
    \label{fig:aws_map}
\end{figure}

>>>>

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.3]{Figures/storhofdi.jpeg}
  \caption[AWS at Stórhöfði in Vestmannaeyjar, in south Iceland.]{AWS at Stórhöfði in Vestmannaeyjar, south of Iceland. The anemometer can be seen as three pins in the upper-central area of the figure just slightly to the left.}
  \label{fig:storhofdi}
\end{figure}

\section{Automatic Weather Station Data}
Automatic weather stations (AWS) have been set up all around the world, including Iceland. AWSs might include a variety of sensors, such as barometers, thermometers, and anemometer along with solar panels (or other power source) and some kind of data logger. IMO provided measurements from AWSs across Iceland. One of the stations used in this study can be seen in Figure \ref{fig:storhofdi}. Not every station had consistently available data. Data from \nStationsMin AWSs were included looking at a period from \startDateVedur and ended in 2023. Of these \nStationsMin stations provided by IMO, \nVedurMin were IMO stations, with the anemometer at 10~m above ground, while the remaining \nVGMin stations were from \href{https://www.vegagerdin.is/}{the Icelandic Road and Coastal Administration (IRCA)}, with the anemometer at 6–7~m above ground \cite{vegagerdin_postur}. The locations of these weather stations are shown in Figure \ref{fig:aws_map}.

@@

\begin{figure}
    \centering
    \includegraphics[width=1.1\textwidth]{Figures/station_elevations.png}
    \caption[Locations of automatic weather stations in Iceland]{Locations of all 412 stations that were looked at in this study. Most of these were from IMO but over a hundred were from IRCA. IMO anemometers are placed at 10 meters above ground, while IRCA ones are placed at 6--7 meters above ground.}
    \label{fig:aws_map}
\end{figure}

### 2025-05-09 Chapters/Chapter3.tex
vinna

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.5]{Figures/errorMap.png}
\caption[MAPE distribution by station]{Mean absolute percentage error (MAPE) at each station, shown as colored circles (dark blue = low error; yellow = high error). The lowest station error is about 7\% at Garðskagaviti and the highest about 35\% at Seljalandsdalur. The model uses a 10 m/s lower bound based on ws$_{15}$.}
\label{fig:errorMap}
\end{figure}

\begin{table}[h]
    \caption[Primary model MAPE by selected stations]{Constant baseline and primary model MAPE for a few selected stations. High MAPE values is seen in fjords near tall mountains in Austfirðir (Fáskrúðsfjörður and Seyðisfjörður), intermediate values at the foothills of mountains (Ingólfsfjall and Kjalarnes), and low values in flatter places (Sandskeið, Þrengsli).}
    \label{table:more_specific_sites}
    \centering
    \begin{tabular}{lcc}
        \toprule
        Station name & \multicolumn{2}{c}{MAPE}\\
         & Baseline & Model\\
        \midrule
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
  \caption[Model result by station]{MAPE for a few selected stations, for a model trained with the whole data set (general training), and for a model trained only with the data of the station itself (site training), with a lower bound of 10 m/s. This investigation was done in an earlier phase of the study, when $f$ instead of ws$_{15}$ was being used as the lower bound cut off, and that leads to some data leakage and a somewhat lower MAPE.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{lccc}
        \toprule
        Station name & n & \multicolumn{2}{c}{MAPE}\\
        & & General training & Site training\\
        \midrule
        Akrafjall & 43,000 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,000 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,000 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,000 & 7.7\% & 7.7\%\\
        Sandbúðir & 19,000 & 6.8\% & 6.4\%\\
        Stórholt & 35,000 & 7.1\% & 29.2\% \\
        Þúfuver & 20,000 & 6.4\% & 6.8\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result looking at closed wind speed intervals]{The MAPE results for different wind speed limit intervals. Here instead of training for all data above a certain threshold of wind speed, training is done only on data between two wind speeds. The percentage variance in gust factor as a function of wind speed increases with decreasing wind speed. Measured wind speed is used for the cutoff and thus have data leakage. These results should thus be somewhat comparable to the last column in Table \ref{table:results}}
    \label{table:closed_intervals}
    \centering
    \begin{tabular}{lcc}
        \toprule
        Interval &  \multicolumn{2}{c}{MAPE}\\
        $[m/s]$ & Without DEM & With DEM \\
        \midrule
        $[5, 10[$ & 17.1\% & 16.4\%\\
        $[10, 15[$ & 14.5\% & 13.0\%\\
        $[15, 20[$ & 15.0\% & 12.0\%\\
        $[20, 25[$ & 15.6 \% & 13.1\%\\
        $[25, 30[$ & 18.4\% & 19.0\%\\
        \bottomrule
    \end{tabular}
\end{table}

\section{Feature importance with Shapley values}
In this section Shapley values (see section \ref{sec:methodology} and \ref{sec:layout}) for the features of the secondary parameter set will be considered. Each feature value in each data row will contribute some share in the final prediction of an NN model, that can be computed by shifting the respective value and observe the change in the model output. This procedure is illustrated in Figure \ref{fig:ShapleyWaterfall}.

\begin{figure}[h]
  \centering
  \includegraphics[scale = 0.6]{Figures/shap_plots/waterfall_plot.png}
  \caption[Shapley contributions for a single observation]{Shapley contributions for a single data row. The figure shows an example of the individual values contributing to a Shapley summary plot. Each point in a summary plot corresponds to one such instance. Several values with the same SHAP value will result in a thicker bulge in the summary. In this specific instance the wind direction ($wd_{15}$) has the highest negative contribution and the temperature has the highest positive contribution. The last three features contribute nothing.}
  \label{fig:ShapleyWaterfall}
\end{figure}

\subsection{Shapley summary plots for all stations}

Figure \ref{fig:ShapleySummary} provides a Shapley summary plot showing the contribution of each feature in the set. Computation of these values is expensive so subset of data is used ($\sim$1000 randomly selected data rows). There is an obvious outlier in the figure, making it difficult to assess the distribution. A second Shapley summary was computed with more data rows ($\sim$10000) and outliers excluded, see the lower panel of the figure. Note that it is not feasible to include the DEM data in this analysis, as they contain far too many variables.

>>>>

\begin{table}[H]
    \caption[Primary model MAPE by selected stations]{Constant baseline and primary model MAPE for a few selected stations. High MAPE values is seen in fjords near tall mountains in Austfirðir (Fáskrúðsfjörður and Seyðisfjörður), intermediate values at the foothills of mountains (Ingólfsfjall and Kjalarnes), and low values in flatter places (Sandskeið, Þrengsli).}
    \label{table:more_specific_sites}
    \centering
    \begin{tabular}{lcc}
        \toprule
        Station name & \multicolumn{2}{c}{MAPE}\\
         & Baseline & Model\\
        \midrule
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

@@

\begin{figure}[H]
  \centering
  \makebox[\textwidth[c]{
    \includegraphics[width=1.1\textwidth]{Figures/errorMap.png}
  }
\caption[MAPE distribution by station]{Mean absolute percentage error (MAPE) at each station, shown as colored circles (dark blue = low error; yellow = high error). The lowest station error is about 7\% at Garðskagaviti and the highest about 35\% at Seljalandsdalur. The model uses a 10 m/s lower bound based on ws$_{15}$.}
\label{fig:errorMap}
\end{figure}

\begin{table}[H]
  \caption[Model result by station]{MAPE for a few selected stations, for a model trained with the whole data set (general training), and for a model trained only with the data of the station itself (site training), with a lower bound of 10 m/s. This investigation was done in an earlier phase of the study, when $f$ instead of ws$_{15}$ was being used as the lower bound cut off, and that leads to some data leakage and a somewhat lower MAPE.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{lccc}
        \toprule
        Station name & n & \multicolumn{2}{c}{MAPE}\\
        & & General training & Site training\\
        \midrule
        Akrafjall & 43,000 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,000 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,000 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,000 & 7.7\% & 7.7\%\\
        Sandbúðir & 19,000 & 6.8\% & 6.4\%\\
        Stórholt & 35,000 & 7.1\% & 29.2\% \\
        Þúfuver & 20,000 & 6.4\% & 6.8\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\section{Feature importance with Shapley values}\label{sec:shapley}
In this section Shapley values (see section \ref{sec:methodology} and \ref{sec:layout}) for the features of the secondary parameter set will be considered. Each feature value in each data row will contribute some share in the final prediction of an NN model, that can be computed by shifting the respective value and observe the change in the model output. This procedure is illustrated in Figure \ref{fig:ShapleyWaterfall}.

\begin{table}
    \caption[Model result looking at closed wind speed intervals]{The MAPE results for different wind speed limit intervals. Here instead of training for all data above a certain threshold of wind speed, training is done only on data between two wind speeds. The percentage variance in gust factor as a function of wind speed increases with decreasing wind speed. Measured wind speed is used for the cutoff and thus have data leakage. These results should thus be somewhat comparable to the last column in Table \ref{table:results}}
    \label{table:closed_intervals}
    \centering
    \begin{tabular}{lcc}
        \toprule
        Interval &  \multicolumn{2}{c}{MAPE}\\
        $[m/s]$ & Without DEM & With DEM \\
        \midrule
        $[5, 10[$ & 17.1\% & 16.4\%\\
        $[10, 15[$ & 14.5\% & 13.0\%\\
        $[15, 20[$ & 15.0\% & 12.0\%\\
        $[20, 25[$ & 15.6 \% & 13.1\%\\
        $[25, 30[$ & 18.4\% & 19.0\%\\
        \bottomrule
    \end{tabular}
\end{table}

\begin{figure}
  \centering
  \includegraphics[scale = 0.6]{Figures/shap_plots/waterfall_plot.png}
  \caption[Shapley contributions for a single observation]{Shapley contributions for a single data row. The figure shows an example of the individual values contributing to a Shapley summary plot. Each point in a summary plot corresponds to one such instance. Several values with the same SHAP value will result in a thicker bulge in the summary. In this specific instance the wind direction ($wd_{15}$) has the highest negative contribution and the temperature has the highest positive contribution. The last three features contribute nothing.}
  \label{fig:ShapleyWaterfall}
\end{figure}

\subsection{Shapley summary plots for all stations}
Figure \ref{fig:ShapleySummary} provides a Shapley summary plot showing the contribution of each feature in the set. Computation of these values is expensive so subset of data is used ($\sim$1000 randomly selected data rows). There is an obvious outlier in the figure, making it difficult to assess the distribution. A second Shapley summary was computed with more data rows ($\sim$10000) and outliers excluded, see the lower panel of the figure. Note that it is not feasible to include the DEM data in this analysis, as they contain far too many variables.

### 2025-05-09 Chapters/Chapter4.tex
vinna

\chapter{Discussion and conclusions}
\label{Chapter5}
As shown in Table \ref{table:results}, the neural-network model achieves a significant reduction in mean absolute percentage error (MAPE) compared to a constant baseline. Table \ref{table:setsOfParams} further demonstrates an improvement over a simple regression model. A statistical test yields $p<0.01$, confirming that these improvements are unlikely to be due to chance. The influence of individual features on model performance is illustrated by Shapley plots (Figure \ref{fig:ShapleySummaryKeflavikurflugvollur} and Appendix \ref{appendix:A}). Although one might expect stability indicators—such as the Richardson number and the Brunt--Väisälä frequency—to be among the strongest predictors of gust variability, they contributed less than other variables. For reanalysis wind speeds above 10~m/s, the network reduces error by about 30\% relative to the regression baseline. The largest gains appear in the 10--25~m/s range (Table \ref{table:closed_intervals}). At lower speeds, percentage errors inflate even when absolute gust--wind differences remain small; at higher speeds, sample sizes become sparse.

@@

This study did not include station coordinates (XY) as direct inputs. Adding location information could improve site-specific accuracy without losing the benefits of training on the full dataset. Similarly, instead of applying a fixed linear interpolation to CARRA and elevation grids, the network could receive raw neighboring grid values and their relative positions, allowing it to learn an optimal interpolation scheme.

Finally, treating the data as a time series may offer further gains. Here, only instantaneous inputs at forecast time were used, partly because CARRA fields are available every three hours and may not capture rapid stability changes. Access to higher-frequency data or the inclusion of temporal history could enable models to exploit evolving atmospheric conditions and further enhance gust-prediction skill.

The power generated by wind turbines increases with the cube of the wind speed \cite{wind_power}. The highest wind gusts in Iceland are around 70 m/s. Knowing the gust factor with half as much error as before can allow better anticipation and thus spare turbines for high wind gusts.

>>>>

\chapter{Discussion and conclusions}
\label{Chapter5}

\section{Main findings}
As shown in Table \ref{table:setsOfParams}, the neural-network model achieves a significant reduction in mean absolute percentage error (MAPE) compared to a constant baseline as well as a simple regression model. Table \ref{table:results} further demonstrates this improvement. For reanalysis wind speeds above 10~m/s, the network reduces error by about 30\% relative to the regression baseline. Table \ref{table:closed_intervals} shows the results of training and model evaluation within closed intervals of the reanalysis wind speed. The best results appear in the 10--25~m/s range. At lower reanlysis wind speeds, percentage errors inflate even when the differences between wind gust and average wind speed remain small and at both at the higher and the lower speeds the base line MAPE is higher.

Although one might expect staility indicators, such as the Richardson number and the Brunt--Väisälä frequency, to be among the strongest predictors of gust variability, they contributed less than other variables. The influence of individual features on model performance is illustrated by Shapley plots (Figure \ref{fig:ShapleySummaryKeflavikurflugvollur} and Appendix \ref{appendix:A}). According the Shapley analysis, stations altitude, mean wind speed and wind direction are among the more important features for predicting the gust factor, and also, to some extent, the Brunt--Väisälä frequency ($N^2$). As explained in Section \ref{sec:shapley} it was infeasible to evaluate the DEM data with Shapley values.

@@

\section{Limitations and future work}
This study did not include station coordinates (X, Y) as direct inputs. Adding location information could improve site-specific accuracy without losing the benefits of training on the full dataset. Similarly, instead of applying a fixed linear interpolation to CARRA and DEM grids, the NN training could receive raw neighboring grid values and their relative positions as part of their input, allowing the model to learn an optimal interpolation scheme.

Finally, treating the data as a time series may offer further gains. Here, only instantaneous inputs at forecast time were used, partly because CARRA fields are available every three hours and may not capture rapid stability changes. Access to higher-frequency data or the inclusion of temporal history could enable models to exploit evolving atmospheric conditions and further enhance gust-prediction skills.

Many papers cited in this thesis suggest that adding a machine learning model as a backend to more traditional numerical atmospheric models could lead to improved performance. The results of this study support this.

\section{Possible applications}
Wind gusts at weather stations are displayed on road signs in Iceland, if wind gusts exceed \href{https://www.vegagerdin.is/vegagerdin/starfsemi/frettir/tvo-ny-vedur-og-upplysingaskilti-i-notkun#}{expected values}. The model could be used to allow such predicitons of wind gusts to be displayed for every section of the road, not just at the stations. In 2023, outside of Reykjavik harbor, a cruise ship nearly \href{https://www.mbl.is/200milur/frettir/2024/04/26/strand_skemmtiferdaskips_kallar_a_erlenda_adstod/}{ran aground} due to an unexpected wind gust. This wind gust of \textasciitilde 26 m/s, pushed the ship into a buoy near the island Viðey. It was only luck that prevented the ship's screws from hitting the buoy. The ship was only 10 meters from the shoals of Viðey, where the water is only 0.4 meters deep. A better prediction of wind gust could have prevented this incidence. Similarly, the model could be used to prevent damage to transportation vehicles, that take on a lot of wind, such as trucks and buses, or to prevent damage to buildings.

The power generated by wind turbines increases with the cube of the wind speed \cite{wind_power}. The highest wind gusts in Iceland are around 70 m/s. Knowing the gust factor accurately would allow the turbine to be stopped in time, thus avoiding damage.

\section{Conclusions}
The neural network model presented in this study demonstrates a significant improvement in predicting wind gusts over a simple regression model, particularly for reanalysis wind speeds above 10 m/s. The model's performance varies by station, and while some stations benefit from the full dataset, others may require more localized training.

### 2025-05-09 MSc.tex
Updated pages and figures.
vinna

\def\thesisyear{2025}       					% Year thesis submitted
\def\thesismonth{June}					% Month thesis submitted, e.g. "December"
\def\thesisauthor{Brynjar Geir Sigurðsson}				% Thesis authoreiningaraðferðinni
\def\thesistitle{Prediction of Gust Factor in Complex Landscape with Neural Networks} % Title of thesis
%\def\thesissubtitle{XXSubtitle is rarely usedXX}		% Subtitle of thesis (optional)
\def\thesisshorttitle{} 	% Optional: if title of thesis is longer than 50 characters, it would not fit on the spine (kjölur) and in this case you, need to provide a short title here for the print shop. Otherwhise: make it empty
\def\thesiscredits{60} 						% Credits awarded for the project
\def\thesissubject{Mechanical Engineering}
\def\thesiskind{M.Sc.}					% E.g. M.Sc. or Ph.D. thesis
\def\thesiskindformal{Magister Scientiarum}			% E.g. Magister Scientiarum
\def\thesisschool{School of Engineering and {Natural Sciences}}		% School
\def\thesisfaculty{Industrial Engineering, Mechanical Engineering and Computer Science}% Faculty name without "Faculty of" (gets added) 
\def\thesisaddress{Dunhagi 5}			        % Faculty or school office address street
\def\thesispostalcode{107}			                % Faculty or school office zip code
\def\thesistelephone{525 4000}					% Office telephone
%\def\thesispublisher{XX}					% Publisher (not used)
\def\thesissupervisors{Kristján Jónasson}					% Names of supervisors (split by \\ if more than one).
\def\thesisnrofsupervisors{1}					% Number of supervisors (to use "Supervisor" vs. "Supervisors")
\def\thesiscommittee{Kristján Jónasson \\ Ólafur Pétur Pálsson \\ Guðrún Nína Petersen}			% Thesis commitee must include the supervisor(s)
\def\thesisexaminer{Halldór Björnsson}				        % Examiner (prófdómari)
\def\thesisISBN{}           					% Thesis ISBN number (keep empty: not used anymore)
\def\thesisprinting{}						% Name of printsthop (keep empty if thesis is never printed)
%\def\thesisprinting{Háskólaprent, Fálkagata 2, 107 Reykjavík}	% Name of printsthop (keep empty if thesis is never printed)
\def\thesislicense{This thesis may not be copied in any form without author permission.} % Set license here (could also be some Creative Commons license)
%\def\thesiskeywords{Keyword1, Keyword2, Keyword3}		% Keywords (not used anywhere, hence commented out)

@@

  \newpage

  \thispagestyle{empty} \mbox{} % This is the inside page of the cover (cover verso) which remains empty

  \newpage

@@

  \newpage

@@

\section*{Útdráttur}
Hvort sem um er að ræða vindmyllugarða, umferðarslys, háspennumöstur, eða aðra hluti sem vindur getur haft áhrif á er spágeta um álag vegna vindsins mikilvæg, til að minnka eða koma í veg fyrir tjón, auka rekstraröryggi, eða bæta framleiðni. Vindhviður eru nokkurra sekúndna aukning á vindi og hviðustuðull mælir hæsta hlutfall milli þriggja sekúndna hviðu og 10 mínútna meðalvinds. Góðar hviðuspár eru mikilvægar. Sögulega hefur verið notast við hefðbundin spálíkön sem reiða sig á töluleg eðlisfræðileg líkön til að spá fyrir um veður, þar með talið hviðustuðul. Á síðustu árum hefur gervigreind þróast mikið og líkön verið þjálfuð til þess að spá fyrir um veður. Svo miklar framfarir hafa orðið að sum líkön geta keppt við hefðbundin veðurlíkön. Markmið þessa verkefnis er reyna að nýta bæði hefðbundin veðurlíkön og gervigreind til þess að meta hviðustuðla. Þetta er gert með því að nota endurgreiningargögn úr hefðbundnu veðurlíkani sem grunngögn í gervigreindarlíkan. Einnig eru notað hæðarlíkan sem lýsir landslagi. Djúpt tauganet er búið til með því að velja breytur úr endurgreiningargögnunum og hæðarlíkaninu og afleiddum breytum. Nokkur líkön af þessu tagi eru svo þjálfuð og niðurstöður bornar saman við einfalt grunnlíkan sem fengið er með línulega aðhvarfi. Niðurstöðurnar lofa mjög góðu um að þessi aðferðafræði geti bætt bæði samtímamat á hviðustuðli og spár um hann umtalsvert. Þetta á ekki síst við í flóknu landslagi.
\vfill

\newpage % Empty page
\setkomafont{section}{\FloatBarrier\Large}
% The first "chapter" which will start on a new page
% Table of contents starts automatically on a right-hand side page ("recto").
% Table of contents, list of figures and tables are automatically generated by the commands below
\hypersetup{pageanchor=true}
\tableofcontents
\listoffigures
\listoftables
\lstlistoflistings

@@

Finally, I would like to thank my parents without whom I would probably never had any want to finish.

\pagenumbering{arabic}
\setcounter{page}{1}

>>>>

\def\thesisyear{2025}       					% Year thesis submitted
\def\thesismonth{May}					% Month thesis submitted, e.g. "December"
\def\thesisauthor{Brynjar Geir Sigurðsson}				% Thesis authoreiningaraðferðinni
\def\thesistitle{Prediction of Gust Factor in \\Complex Landscape with Neural Networks} % Title of thesis
%\def\thesissubtitle{XXSubtitle is rarely usedXX}		% Subtitle of thesis (optional)
\def\thesisshorttitle{} 	% Optional: if title of thesis is longer than 50 characters, it would not fit on the spine (kjölur) and in this case you, need to provide a short title here for the print shop. Otherwhise: make it empty
\def\thesiscredits{60} 						% Credits awarded for the project
\def\thesissubject{Mechanical Engineering}
\def\thesiskind{M.Sc.}					% E.g. M.Sc. or Ph.D. thesis
\def\thesiskindformal{Magister Scientiarum}			% E.g. Magister Scientiarum
\def\thesisschool{School of Engineering and {Natural Sciences}}		% School
\def\thesisfaculty{Industrial Engineering, Mechanical Engineering and Computer Science}% Faculty name without "Faculty of" (gets added) 
\def\thesisaddress{Dunhagi 5}			        % Faculty or school office address street
\def\thesispostalcode{107}			                % Faculty or school office zip code
\def\thesistelephone{525 4000}					% Office telephone
%\def\thesispublisher{XX}					% Publisher (not used)
\def\thesissupervisors{Kristján Jónasson}					% Names of supervisors (split by \\ if more than one).
\def\thesisnrofsupervisors{1}					% Number of supervisors (to use "Supervisor" vs. "Supervisors")
\def\thesiscommittee{Kristján Jónasson \\ Ólafur Pétur Pálsson \\ Guðrún Nína Petersen}			% Thesis commitee must include the supervisor(s)
\def\thesisexaminer{Halldór Björnsson}				        % Examiner (prófdómari)
\def\thesisISBN{}           					% Thesis ISBN number (keep empty: not used anymore)
\def\thesisprinting{}						% Name of printsthop (keep empty if thesis is never printed)
%\def\thesisprinting{Háskólaprent, Fálkagata 2, 107 Reykjavík}	% Name of printsthop (keep empty if thesis is never printed)
\def\thesislicense{This thesis may not be copied in any form without author permission.} % Set license here (could also be some Creative Commons license)
%\def\thesiskeywords{Keyword1, Keyword2, Keyword3}		% Keywords (not used anywhere, hence commented out)

@@

  %\newpage

  %\thispagestyle{empty} \mbox{} % This is the inside page of the cover (cover verso) which remains empty

  %\newpage

@@

  %\newpage

@@

\section*{Útdráttur}
Hvort sem um er að ræða vindmyllugarða, umferðarslys, háspennumöstur, eða aðra hluti sem vindur getur haft áhrif á er spágeta um álag vegna vindsins mikilvæg, til að minnka eða koma í veg fyrir tjón, auka rekstraröryggi, eða bæta framleiðni. Vindhviður eru nokkurra sekúndna aukning á vindi og hviðustuðull mælir hæsta hlutfall milli þriggja sekúndna hviðu og 10 mínútna meðalvinds. Góðar hviðuspár eru mikilvægar. Sögulega hefur verið notast við hefðbundin spálíkön sem reiða sig á töluleg eðlisfræðileg líkön til að spá fyrir um veður, þar með talið hviðustuðul. Á síðustu árum hefur gervigreind þróast mikið og líkön verið þjálfuð til þess að spá fyrir um veður. Svo miklar framfarir hafa orðið að sum líkön geta keppt við hefðbundin veðurlíkön. Markmið þessa verkefnis er reyna að nýta bæði hefðbundin veðurlíkön og gervigreind til þess að meta hviðustuðla. Þetta er gert með því að nota endurgreiningargögn úr hefðbundnu veðurlíkani sem grunngögn í gervigreindarlíkan. Einnig eru notað hæðarlíkan sem lýsir landslagi. Djúpt tauganet er búið til með því að velja breytur úr endurgreiningargögnunum og hæðarlíkaninu og afleiddum breytum. Nokkur líkön af þessu tagi eru svo þjálfuð og niðurstöður bornar saman við einfalt grunnlíkan sem fengið er með línulega aðhvarfi. Niðurstöðurnar lofa mjög góðu um að þessi aðferðafræði geti bætt bæði samtímamat á hviðustuðli og spár um hann umtalsvert. Þetta á ekki síst við í flóknu landslagi.
%\vfill

%\newpage % Empty page
\setkomafont{section}{\FloatBarrier\Large}
% The first "chapter" which will start on a new page
% Table of contents starts automatically on a right-hand side page ("recto").
% Table of contents, list of figures and tables are automatically generated by the commands below
\hypersetup{pageanchor=true}
\tableofcontents
\listoffigures
\listoftables
\lstlistoflistings

@@

Finally, I would like to thank my parents without whom I would probably never had any want to finish.
\newpage
\pagenumbering{arabic}
\setcounter{page}{1}
